---
title: "Blockchain Real Estate Rental"
author: "Indrek Värva (B26236)"
date: "5/13/2018"
output: html_document
---

```{r setup, include=TRUE}
library(ggplot2)
library(scales)
library(gridExtra)
options(scipen=10000)
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, fig.width=12)
```

Read the data in.

```{r}
v1deploymentGas = read.csv(file=file.path("data", "v1", "deployment-gas.csv"), header = TRUE, sep = ",");
v1offerGas = read.csv(file=file.path("data", "v1", "offer-gas.csv"), header = TRUE, sep = ",")
v1scenarioGas = read.csv(file=file.path("data", "v1", "scenario-gas.csv"), header = TRUE, sep = ",")
v1scenarioRequests = read.csv(file=file.path("data", "v1", "scenario-requests.csv"), header = TRUE, sep = ",")
v1scenarioTimer = read.csv(file=file.path("data", "v1", "scenario-timer.csv"), header = TRUE, sep = ",")
v1geoScenarioGas = read.csv(file=file.path("data", "v1", "geosearch-exclusive-scenario-gas.csv"), header = TRUE, sep = ",")
v1geoScenarioRequests = read.csv(file=file.path("data", "v1", "geosearch-exclusive-scenario-requests.csv"), header = TRUE, sep = ",")
v1geoScenarioTimer = read.csv(file=file.path("data", "v1", "geosearch-exclusive-scenario-timer.csv"), header = TRUE, sep = ",")

v2deploymentGas = read.csv(file=file.path("data", "v2", "deployment-gas.csv"), header = TRUE, sep = ",");
v2offerGas = read.csv(file=file.path("data", "v2", "offer-gas.csv"), header = TRUE, sep = ",")
v2scenarioGas = read.csv(file=file.path("data", "v2", "scenario-gas.csv"), header = TRUE, sep = ",")
v2scenarioRequests = read.csv(file=file.path("data", "v2", "scenario-requests.csv"), header = TRUE, sep = ",")
v2scenarioTimer = read.csv(file=file.path("data", "v2", "scenario-timer.csv"), header = TRUE, sep = ",")
v2geoScenarioGas = read.csv(file=file.path("data", "v2", "geosearch-exclusive-scenario-gas.csv"), header = TRUE, sep = ",")
v2geoScenarioRequests = read.csv(file=file.path("data", "v2", "geosearch-exclusive-scenario-requests.csv"), header = TRUE, sep = ",")
v2geoScenarioTimer = read.csv(file=file.path("data", "v2", "geosearch-exclusive-scenario-timer.csv"), header = TRUE, sep = ",")

v3deploymentGas = read.csv(file=file.path("data", "v3", "deployment-gas.csv"), header = TRUE, sep = ",");
v3offerGas = read.csv(file=file.path("data", "v3", "offer-gas.csv"), header = TRUE, sep = ",")
v3scenarioGas = read.csv(file=file.path("data", "v3", "scenario-gas.csv"), header = TRUE, sep = ",")
v3scenarioRequests = read.csv(file=file.path("data", "v3", "scenario-requests.csv"), header = TRUE, sep = ",")
v3scenarioTimer = read.csv(file=file.path("data", "v3", "scenario-timer.csv"), header = TRUE, sep = ",")
v3geoScenarioGas = read.csv(file=file.path("data", "v3", "geosearch-exclusive-scenario-gas.csv"), header = TRUE, sep = ",")
v3geoScenarioRequests = read.csv(file=file.path("data", "v3", "geosearch-exclusive-scenario-requests.csv"), header = TRUE, sep = ",")
v3geoScenarioTimer = read.csv(file=file.path("data", "v3", "geosearch-exclusive-scenario-timer.csv"), header = TRUE, sep = ",")
```

Reshape the data to a necessary form.

```{r}
prepareData <- function(df,cls) {
  m = ncol(df)
  enlistments = rep(as.factor(as.numeric(rownames(df))-1), each=m)
  df <- t(df)
  n = ncol(df)
  steps = rep(rownames(df), n)
  res = data.frame(run = enlistments, step = steps, value = as.vector(as.matrix(df)))
  res$variant = cls
  return(res)
}

vv1deploymentGas = prepareData(v1deploymentGas, "v1");
vv1offerGas = prepareData(v1offerGas, "v1");
vv1scenarioGas = prepareData(v1scenarioGas, "v1");
vv1scenarioRequests = prepareData(v1scenarioRequests, "v1");
vv1geoScenarioGas = prepareData(v1geoScenarioGas, "v1");
vv1geoScenarioRequests = prepareData(v1geoScenarioRequests, "v1");

vv2deploymentGas = prepareData(v2deploymentGas, "v2");
vv2offerGas = prepareData(v2offerGas, "v2");
vv2scenarioGas = prepareData(v2scenarioGas, "v2");
vv2scenarioRequests = prepareData(v2scenarioRequests, "v2");
vv2geoScenarioGas = prepareData(v2geoScenarioGas, "v2");
vv2geoScenarioRequests = prepareData(v2geoScenarioRequests, "v2");

vv3deploymentGas = prepareData(v3deploymentGas, "v3");
vv3offerGas = prepareData(v3offerGas, "v3");
vv3scenarioGas = prepareData(v3scenarioGas, "v3");
vv3scenarioRequests = prepareData(v3scenarioRequests, "v3");
vv3geoScenarioGas = prepareData(v3geoScenarioGas, "v3");
vv3geoScenarioRequests = prepareData(v3geoScenarioRequests, "v3");

# vertical merge the variants
deploymentGas = rbind(vv1deploymentGas, vv2deploymentGas, vv3deploymentGas);
offerGas = rbind(vv1offerGas, vv2offerGas, vv3offerGas);
scenarioGas = rbind(vv1scenarioGas, vv2scenarioGas, vv3scenarioGas);
scenarioRequests = rbind(vv1scenarioRequests, vv2scenarioRequests, vv3scenarioRequests);
geoScenarioGas = rbind(vv1geoScenarioGas, vv2geoScenarioGas, vv3geoScenarioGas);
geoScenarioRequests = rbind(vv1geoScenarioRequests, vv2geoScenarioRequests, vv3geoScenarioRequests);

```


Deployment cost

```{r}
# deploymentGas, offerGas, scenarioGas, scenarioRequests, geoScenarioGas, geoScenarioRequests


deploymentGas$step = as.character(deploymentGas$step)
deploymentGas$step[deploymentGas$step == 'trigonometry'] <- 'Trigonometry library'
deploymentGas$step[deploymentGas$step == 'geodistance'] <- 'Geodistance library'
deploymentGas$step[deploymentGas$step == 'enlistmentregistry'] <- 'Enlistment Registry'
deploymentGas$step[deploymentGas$step == 'enlistment'] <- 'Enlistment'
deploymentGas$step <- factor(deploymentGas$step, levels=c("Trigonometry library", "Geodistance library", "Enlistment Registry", "Enlistment"))

deploymentGas$variant = as.character(deploymentGas$variant)
deploymentGas$variant[deploymentGas$variant == 'v1'] <- 'Baseline'
deploymentGas$variant[deploymentGas$variant == 'v2'] <- 'Data migration'
deploymentGas$variant[deploymentGas$variant == 'v3'] <- 'Operation migration'

# Final plot 2
ggplot(data = deploymentGas, aes(x=variant, y=value, fill=step)) +
  geom_bar(stat="identity", color="black") +
  ylab("Gas") +
  xlab("Variant") +
  guides(fill=guide_legend(title="Smart Contract")) +
  theme_bw(base_size=22) +
  theme(
    text = element_text(size=22),
    plot.title = element_blank(),
    axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
    legend.key.size = unit(2, 'lines')
    ) + 
  scale_y_continuous(breaks = round(seq(0, 8000000, by = 500000),2))

# Final plot 1
omitEnlistmentDeployment <- subset(deploymentGas, step != "Enlistment")

# to explicitly show on the plot that variant 1 has no singleton gas usage, add a row with value 0
zeroValueRow<-data.frame("0","Enlistment Registry",0,"Baseline")
names(zeroValueRow)<-c("run", "step","value","variant")

omitEnlistmentDeploymentVariant1Added <- rbind(omitEnlistmentDeployment,zeroValueRow)
ggplot(data = omitEnlistmentDeploymentVariant1Added, aes(x=variant, y=value, fill=step)) +
  geom_bar(stat="identity", color="black") +
  ggtitle('Comparison of System Singleton Smart Contract Deployment Gas Usage') +
  ylab("Gas") +
  xlab("Variant") +
  guides(fill=guide_legend(title="Smart Contract")) +
  theme_bw(base_size=22) +
  theme(
    text = element_text(size=22),
    plot.title = element_blank(),
    axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
    legend.key.size = unit(2, 'lines')
    ) +
  scale_x_discrete(drop=FALSE) +
  scale_y_continuous(breaks = round(seq(0, 8000000, by = 100000),2))

```

RESULT INTERPRETATION

1. Contract deployment costs for all versions.

In terms of smart contract deployment costs, variation 3 is the most demanding. When comparing the gas consumption for singleton contract deployments required for the variations to operate, v3 deploys a total of 3 contracts. Meanwhile, v2 only requires 1 singleton contract and v1 operates without any singletons. However, for the case of multiple-instance contract to manage rental intermediation subprocesses, the gas consumption gets bigger with every iteration.

```{r}
# progressive ops: 1, 5, 9, 10, 13, 14, 15, 16
includedSteps = c('step1', 'step5', 'step9', 'step10', 'step13', 'step14', 'step15', 'step16')
N = 10

pData = subset(scenarioGas, step %in% includedSteps & run == '10')
pData$variant = as.character(pData$variant)
pData$variant[pData$variant == 'v1'] <- 'Baseline'
pData$variant[pData$variant == 'v2'] <- 'Data migration'
pData$variant[pData$variant == 'v3'] <- 'Operation migration'

# subtract enlistment deployment cost from step1
v1eCost = 4574231
v2eCost = 5698313
v3eCost = 5599812

pData$step <- factor(pData$step, levels=c("step1","step2","step3","step4","step5","step6","step7","step8","step9","step10","step11","step12","step13","step14","step15","step16"))

# final plot 3
ggplot(data = pData, aes(x=step, y=value, fill=variant)) +
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  ylab("Gas") +
  xlab("Step") +
  guides(fill=guide_legend(title="Variant")) +
  theme_bw(base_size=22) +
  theme(
    text = element_text(size=22),
    plot.title = element_blank(),
    axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
    legend.key.size = unit(2, 'lines')
    ) +
  scale_x_discrete(drop=TRUE) +
  scale_y_continuous(breaks = round(seq(0, 8000000, by = 500000),2))

prData <- pData
prData$value[prData$step == 'step1'] = prData$value[prData$step == 'step1'] - c(v1eCost, v2eCost, v3eCost)

# final plot 4
ggplot(data = prData, aes(x=step, y=value, fill=variant)) +
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  ylab("Gas") + 
  xlab("Step") + 
  guides(fill=guide_legend(title="Variant")) + 
  theme_bw(base_size=22) +
  theme(
    text = element_text(size=22),
    axis.title.x = element_text(margin = margin(t = 30, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 30, b = 0, l = 0)),
    legend.key.size = unit(2, 'lines')
    ) + 
  scale_x_discrete(drop=TRUE) +
  scale_y_continuous(breaks = round(seq(0, 8000000, by = 50000),2))

# sum(subset(prData, variant=='Baseline')$value)

```

```{r}
# prepare data for arbitrary prior data analysis
# progressive ops: 1, 5, 9, 10, 13, 14, 15, 16
# change of gas usage for arbitrary N

pfData = subset(scenarioGas, step %in% includedSteps)
pfData$variant = as.character(pfData$variant)
pfData$variant[pfData$variant == 'v1'] <- 'Baseline'
pfData$variant[pfData$variant == 'v2'] <- 'Data migration'
pfData$variant[pfData$variant == 'v3'] <- 'Operation migration'
pfData$step <- factor(pfData$step, levels=c("step1","step2","step3","step4","step5","step6","step7","step8","step9","step10","step11","step12","step13","step14","step15","step16"))

pf1Data = subset(pfData, step == 'step1')
pf5Data = subset(pfData, step == 'step5')
pf9Data = subset(pfData, step == 'step9')
pf10Data = subset(pfData, step == 'step10')
pf13Data = subset(pfData, step == 'step13')
pf14Data = subset(pfData, step == 'step14')
pf15Data = subset(pfData, step == 'step15')
pf16Data = subset(pfData, step == 'step16')
```


```{r}


#legend example
ggplot(pf15Data, aes(x=run,y=value,group=variant)) + 
  geom_point(aes(color=variant, fill=variant, shape=variant, size=variant), color="black", shape=21, alpha=1) + 
  ggtitle('Step 15 Gas Usage for 0..31 Prior Enlistments\nin the Registry (N)') +
  ylab("Gas [wei]") + xlab("N") +
  guides(fill=guide_legend(title="Variant", override.aes = list(size=4))) +
  guides(size=FALSE) +
  theme(
    text = element_text(size=28),
    axis.title.x = element_text(margin = margin(t = 30, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 30, b = 0, l = 0)),
    legend.text = element_text(size=28),
    legend.key.size = unit(2.5, 'lines')
    ) +
  scale_x_discrete(drop=TRUE, breaks = round(seq(0, 31, by = 8),2)) +
  scale_y_continuous(expand = c(0.3, 0.3)) +
  scale_size_manual(values=c(4,5,2)) +
  scale_fill_manual(values=c("#D55E00", "#0072B2", "#F0E442")) 


getGasPlot <- function(df, stepNr) {
  return(
    ggplot(df, aes(x=run,y=value,group=variant)) +
      geom_point(aes(color=variant, fill=variant, shape=variant, size=variant), shape=21, alpha=1) +
      geom_line(alpha=0.6) +
      guides(size=FALSE) +
      guides(fill=FALSE) +
      guides(color=FALSE) +
      guides(shape=FALSE) +
      xlab("N") +
      ggtitle(paste("Step ", stepNr)) +
      theme_bw(base_size=14) +
      theme(
        text = element_text(size=18),
        axis.title.x = element_text(size=14),
        axis.title.y = element_blank(),
        axis.text = element_text(size=13),
        plot.title = element_text(size=16, hjust = 0.5),
        legend.position = "none"
        ) +
      scale_y_continuous(expand = c(0.3, 0.3)) +
      scale_x_discrete(drop=TRUE, breaks = round(seq(0, 31, by = 8),2)) +
      scale_size_manual(values=c(6,4,2)) +
      scale_fill_manual(values=c("#D55E00", "#0072B2", "#F0E442"))
  );
}

s1p <- getGasPlot(pf1Data, "1")
s5p <- getGasPlot(pf5Data, "5")
s9p <- getGasPlot(pf9Data, "9")
s10p <- getGasPlot(pf10Data, "10")
s13p <- getGasPlot(pf13Data, "13")
s14p <- getGasPlot(pf14Data, "14")
s15p <- getGasPlot(pf15Data, "15")
s16p <- getGasPlot(pf16Data, "16")
# final plot
grid.arrange(s1p, s5p, s9p, s10p, s13p, s14p, s15p, s16p, nrow=3)
```


Request count analysis

```{r}
includedSteps = c('step1', 'step5', 'step9', 'step10', 'step13', 'step14', 'step15', 'step16')
rreqData = subset(geoScenarioRequests, step %in% includedSteps)
rreqData$variant = as.character(rreqData$variant)
rreqData$variant[rreqData$variant == 'v1'] <- 'Baseline'
rreqData$variant[rreqData$variant == 'v2'] <- 'Data migration'
rreqData$variant[rreqData$variant == 'v3'] <- 'Operation migration'
rreqData$step <- factor(rreqData$step, levels=c("step1","step2","step3","step4","step5","step6","step7","step8","step9","step10","step11","step12","step13","step14","step15","step16"))

rreq1Data = subset(rreqData, step == 'step1')
rreq5Data = subset(rreqData, step == 'step5')
rreq9Data = subset(rreqData, step == 'step9')
rreq10Data = subset(rreqData, step == 'step10')
rreq13Data = subset(rreqData, step == 'step13')
rreq14Data = subset(rreqData, step == 'step14')
rreq15Data = subset(rreqData, step == 'step15')
rreq16Data = subset(rreqData, step == 'step16')

requestBreaks <- function() scale_y_continuous(breaks=c(0,1,2,3), expand = c(0.5, 0.5))

r1p <- getGasPlot(rreq1Data, "1") + requestBreaks()
r5p <- getGasPlot(rreq5Data, "5") + requestBreaks()
r9p <- getGasPlot(rreq9Data, "9") + requestBreaks()
r10p <- getGasPlot(rreq10Data, "10") + requestBreaks()
r13p <- getGasPlot(rreq13Data, "13") + requestBreaks()
r14p <- getGasPlot(rreq14Data, "14") + requestBreaks()
r15p <- getGasPlot(rreq15Data, "15") + requestBreaks()
r16p <- getGasPlot(rreq16Data, "16") + requestBreaks()
# final plot
grid.arrange(r1p, r5p, r9p, r10p, r13p, r14p, r15p, r16p, nrow=3)

```


```{r}
# Total mean execution cost (write)
includedSteps = c('step1', 'step5', 'step9', 'step10', 'step13', 'step14', 'step15', 'step16')
# calculate scenario mean for write steps
v1WriteStepsMeans = prepareData(as.data.frame.list((colMeans(v1geoScenarioGas))[includedSteps]), "Baseline")
v1meanCost = sum(v1WriteStepsMeans$value)

v2WriteStepsMeans = prepareData(as.data.frame.list((colMeans(v2geoScenarioGas))[includedSteps]), "Data migration")
v2meanCost = sum(v2WriteStepsMeans$value)

v3WriteStepsMeans = prepareData(as.data.frame.list((colMeans(v3geoScenarioGas))[includedSteps]), "Operation migration")
v3meanCost = sum(v3WriteStepsMeans$value)

meanWriteCosts = rbind(v1WriteStepsMeans, v2WriteStepsMeans, v3WriteStepsMeans)

ggplot(meanWriteCosts, aes(x=step,y=value,group=variant)) +
      geom_point(aes(color=variant, fill=variant, shape=variant, size=variant), shape=21, alpha=1) +
      guides(size=FALSE) +
      guides(fill=FALSE) +
      guides(color=FALSE) +
      guides(shape=FALSE) +
      xlab("N") +
      theme_bw(base_size=14) +
      theme(
        text = element_text(size=18),
        axis.title.x = element_text(size=14),
        axis.title.y = element_blank(),
        axis.text = element_text(size=13),
        plot.title = element_text(size=16, hjust = 0.5),
        legend.position = "none"
        ) +
      scale_y_continuous(expand = c(0.3, 0.3)) +
      scale_size_manual(values=c(6,4,2)) +
      scale_fill_manual(values=c("#D55E00", "#0072B2", "#F0E442"));
```




```{r}
# progressive ops: 2, 3, 4, 6, 7, 8, 11, 12
aIncludedSteps = c('step2', 'step3', 'step4', 'step6', 'step7', 'step8', 'step11', 'step12')
N = 10

oData = subset(geoScenarioGas, step %in% aIncludedSteps & run == '10')
oData$variant = as.character(oData$variant)
oData$variant[oData$variant == 'v1'] <- 'Baseline'
oData$variant[oData$variant == 'v2'] <- 'Data migration'
oData$variant[oData$variant == 'v3'] <- 'Operation migration'


oData$step <- factor(oData$step, levels=c("step1","step2","step3","step4","step5","step6","step7","step8","step9","step10","step11","step12","step13","step14","step15","step16"))

# final plot
ggplot(data = oData, aes(x=step, y=value, fill=variant)) +
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  ylab("Gas") +
  xlab("Step") +
  guides(fill=guide_legend(title="Variant")) +
  theme_bw(base_size=22) +
  theme(
    text = element_text(size=22),
    plot.title = element_blank(),
    axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
    legend.key.size = unit(2, 'lines')
    ) +
  scale_x_discrete(drop=TRUE) +
  scale_y_continuous(breaks = round(seq(0, 8000000, by = 100000),2))

```

Data preparation for arbitrary prior enlistments analysis.
```{r}

# assistive ops: 2, 3, 4, 6, 7, 8, 11, 12

asData = subset(geoScenarioGas, step %in% aIncludedSteps)
asData$variant = as.character(asData$variant)
asData$variant[asData$variant == 'v1'] <- 'Baseline'
asData$variant[asData$variant == 'v2'] <- 'Data migration'
asData$variant[asData$variant == 'v3'] <- 'Operation migration'
asData$step <- factor(asData$step, levels=c("step1","step2","step3","step4","step5","step6","step7","step8","step9","step10","step11","step12","step13","step14","step15","step16"))

as2Data = subset(asData, step == 'step2')
as3Data = subset(asData, step == 'step3')
as4Data = subset(asData, step == 'step4')
as6Data = subset(asData, step == 'step6')
as7Data = subset(asData, step == 'step7')
as8Data = subset(asData, step == 'step8')
as11Data = subset(asData, step == 'step11')
as12Data = subset(asData, step == 'step12')

```


```{r}
getAsGasPlot <- function(df, stepNr) {
  return(
    ggplot(df, aes(x=run,y=value,group=variant)) +
      geom_point(aes(color=variant, fill=variant, shape=variant, size=variant), shape=21, alpha=1) +
      geom_line(alpha=0.6) +
      guides(size=FALSE) +
      guides(fill=FALSE) +
      guides(color=FALSE) +
      guides(shape=FALSE) +
      xlab("N") +
      ggtitle(paste("Step ", stepNr)) +
      theme_bw(base_size=14) +
      theme(
        text = element_text(size=18),
        axis.title.x = element_text(size=14),
        axis.title.y = element_blank(),
        axis.text = element_text(size=13),
        plot.title = element_text(size=16, hjust = 0.5),
        legend.position = "none"
        ) +
      scale_y_continuous(expand = c(0.3, 0.3)) +
      scale_x_discrete(drop=TRUE, breaks = round(seq(0, 31, by = 8),2)) +
      scale_size_manual(values=c(6,4,2)) +
      scale_fill_manual(values=c("#D55E00", "#0072B2", "#F0E442"))
  );
}

# assistive ops: 2, 3, 4, 6, 7, 8, 11, 12

a2p <- getAsGasPlot(as2Data, "2")
a3p <- getAsGasPlot(as3Data, "3")
a4p <- getAsGasPlot(as4Data, "4")
a6p <- getAsGasPlot(as6Data, "6")
a7p <- getAsGasPlot(as7Data, "7")
a8p <- getAsGasPlot(as8Data, "8")
a11p <- getAsGasPlot(as11Data, "11")
a12p <- getAsGasPlot(as12Data, "12")
# final plot
grid.arrange(a2p, a3p, a4p, a6p, a7p, a8p, a11p, a12p, nrow=3)
```

```{r}
# assistive ops: 2, 3, 4, 6, 7, 8, 11, 12
readIncludedSteps = c('step2', 'step3', 'step4', 'step6', 'step7', 'step8', 'step11', 'step12')
xData = subset(geoScenarioRequests, step %in% readIncludedSteps)
xData$variant = as.character(xData$variant)
xData$variant[xData$variant == 'v1'] <- 'Baseline'
xData$variant[xData$variant == 'v2'] <- 'Data migration'
xData$variant[xData$variant == 'v3'] <- 'Operation migration'
xData$step <- factor(xData$step, levels=c("step1","step2","step3","step4","step5","step6","step7","step8","step9","step10","step11","step12","step13","step14","step15","step16"))

x2Data = subset(xData, step == 'step2')
x3Data = subset(xData, step == 'step3')
x4Data = subset(xData, step == 'step4')
x6Data = subset(xData, step == 'step6')
x7Data = subset(xData, step == 'step7')
x8Data = subset(xData, step == 'step8')
x11Data = subset(xData, step == 'step11')
x12Data = subset(xData, step == 'step12')

xRequestBreaks <- function() scale_y_continuous(breaks=c(0,1,2,3), expand = c(0.5, 0.5))

x2p <- getGasPlot(x2Data, "2")
x3p <- getGasPlot(x3Data, "3")
x4p <- getGasPlot(x4Data, "4") + scale_y_continuous(breaks=c(0,1,2,3), expand = c(0.5, 0.5))
x6p <- getGasPlot(x6Data, "6")
x7p <- getGasPlot(x7Data, "7") + scale_y_continuous(breaks=c(3,4,5,6), expand = c(0.5, 0.5))
x8p <- getGasPlot(x8Data, "8") + scale_y_continuous(breaks=c(0,1,2,3), expand = c(0.5, 0.5))
x11p <- getGasPlot(x11Data, "11")
x12p <- getGasPlot(x12Data, "12") + scale_y_continuous(breaks=c(3,4,5,6), expand = c(0.5, 0.5))
# final plot
grid.arrange(x2p, x3p, x4p, x6p, x7p, x8p, x11p, x12p, nrow=3)
```

Mean calculation (read)

```{r}
# assistive ops: 2, 3, 4, 6, 7, 8, 11, 12
readIncludedSteps = c('step2', 'step3', 'step4', 'step6', 'step7', 'step8', 'step11', 'step12')
# calculate scenario mean for write steps
v1ReadStepsMeans = prepareData(as.data.frame.list((colMeans(v1geoScenarioGas))[readIncludedSteps]), "Baseline")
v1meanReadCost = sum(v1ReadStepsMeans$value)

v2ReadStepsMeans = prepareData(as.data.frame.list((colMeans(v2geoScenarioGas))[readIncludedSteps]), "Data migration")
v2meanReadCost = sum(v2ReadStepsMeans$value)

v3ReadStepsMeans = prepareData(as.data.frame.list((colMeans(v3geoScenarioGas))[readIncludedSteps]), "Operation migration")
v3meanReadCost = sum(v3ReadStepsMeans$value)
```


